---
title: "Comparing panel estimators with cfcompare"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Comparing panel estimators with cfcompare}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  # The estimator chunks below run a full TROP cross-validation + jackknife.
  # During R CMD build/check the vignette is knitted from scratch (twice:
  # once at build, once when check re-builds vignette outputs), so leaving
  # these live makes every check dominated by the vignette. We therefore run
  # them ONLY when explicitly opted in. To render with live output locally:
  #   Sys.setenv(CFCOMPARE_BUILD_VIGNETTE = "true")
  #   devtools::build_vignettes()   # or pkgdown::build_site()
  eval = nzchar(Sys.getenv("CFCOMPARE_BUILD_VIGNETTE")) &&
    requireNamespace("ggplot2", quietly = TRUE)
)
```

## The problem

You have a binary-treatment panel and several plausible estimators: classic
difference-in-differences (DID), synthetic difference-in-differences (SDID),
synthetic control (SC), matrix completion (MC), and the Triply RObust Panel
(TROP) estimator of Athey, Imbens, Qu & Viviano (2025). Each lives in its own
package with its own data format, its own notion of a "treatment effect" and its
own plot. Comparing them on the *same* dataset means re-shaping data five times
and reconciling five different output objects by hand.

`cfcompare` does that reconciliation for you. You pass one long-format data frame and
a vector of method names; you get back one tidy table of ATTs and one plot.

```{r setup}
library(cfcompare)
```

## A simulated panel

`sim_panel()` draws from a low-rank factor model
`Y_it(0) = alpha_i + beta_t + L_it + eps_it` and adds a constant treatment effect
to treated cells, so we know the ground truth.

```{r sim}
df <- sim_panel(N = 16, T = 9, n_treated = 3, t0 = 7, rank = 2,
                att = 2, noise = 1, seed = 1)
head(df)
```

The columns are the unit id (`id`), time (`t`), observed outcome (`y`), the 0/1
treatment indicator (`w`), and the oracle untreated outcome (`y0`, for
evaluation only). The true ATT here is `2`.

## Comparing estimators

```{r compare}
cmp <- panel_compare(
  df,
  outcome = "y", treatment = "w", unit = "id", time = "t",
  methods = c("DID", "MC", "TROP"),
  se = "jackknife",
  control = trop_control(n_cv_cells = 12L, cv_cycles = 1L)
)
cmp
```

The `att` element is a `cf_att_tbl`: one row per method, all on the same
schema.

```{r att-table}
cmp$att
```

Every estimate carries the number of treated cells and units, the estimated rank
of the low-rank component `L` (for the native engines), and a `note` column that
records anything that was skipped.

## Plots

`autoplot()` on the comparison gives a forest plot of the ATTs with their
confidence intervals:

```{r forest, fig.width = 6, fig.height = 3}
autoplot(cmp)
```

`plot_counterfactual()` overlays the observed treated-average path against each
estimator's predicted untreated path, so you can see *where* the methods
disagree, not just by how much:

```{r cf, fig.width = 6, fig.height = 3.5}
plot_counterfactual(cmp)
```

## One estimator at a time

`trop()` is the workhorse. Called on its own it cross-validates the three
penalties — time-weight decay, unit-weight decay, and the nuclear-norm penalty
on `L` — by leave-one-out on the control cells.

```{r trop-fit}
fit <- trop(df, "y", "w", "id", "t", se = "jackknife",
            control = trop_control(n_cv_cells = 12L, cv_cycles = 1L))
fit
```

The fitted object exposes the per-cell effects, the estimated untreated-outcome
matrix, and the selected penalties:

```{r trop-internals}
head(fit$tau_cells)
fit$lambda
```

### Recovering special cases

The TROP program nests DID and matrix completion. Fixing the penalties by hand
matches them numerically, which is useful as a sanity check:

```{r special}
did <- trop(df, "y", "w", "id", "t",
            lambda = list(time = 0, unit = 0, nn = Inf))
mc  <- trop(df, "y", "w", "id", "t",
            lambda = list(time = 0, unit = 0, nn = 5))
c(DID = did$att, MC = mc$att, TROP = fit$att)
```

With `lambda_nn = Inf` and uniform weights the low-rank term vanishes and the fit
is ordinary two-way fixed effects; with a finite `lambda_nn` and uniform weights
it is matrix completion; letting cross-validation pick the unit and time weights
gives the full TROP estimator.

## Bringing your own results

If you already ran an estimator elsewhere — say `synthdid` directly — coerce it
onto the shared schema with `as_att()` and combine it with native fits:

```{r byo, eval = FALSE}
library(synthdid)
sd <- synthdid_estimate(Y, N0, T0)

combined <- as_att(list(
  fit,
  as_att(sd, method = "SDID", outcome = "y")
))
autoplot(combined)
```

## Designs and inference

The native engines (DID, MC, TROP) accept **block**, **staggered**, and
**non-absorbing** binary treatments: `treatment` is simply the 0/1 indicator of
active treatment in each cell. `SDID` and `SC` are defined for block designs only
and are skipped, with a note, on other designs.

Standard errors from the native engines are **jackknife**
(leave-one-treated-unit-out) when there are at least two treated units, and
**placebo** when there is a single treated unit. These are practical
approximations; see Section 6 of the paper for the estimator's formal inference
theory.

## Performance

The native solver runs on base R by default, but two settings speed it up on
larger panels (both via `trop_control()`):

* **`svd`** — the soft-impute step uses a **truncated SVD by default**
  (`svd = "truncated"`), computing only the leading singular triplets with
  `RSpectra` when it is installed (and falling back to the full base-R `svd()`
  otherwise). The TROP paper explicitly permits a truncated decomposition for the
  low-rank step, and it matches the full SVD to numerical tolerance. Pass
  `svd = "full"` to force the exact full decomposition; that is what the README's
  numerical-agreement checks against the official Python package use, so the
  comparison is exact rather than up to truncation tolerance.
* **`workers`** — the bootstrap / jackknife / placebo standard errors, the
  cross-validation cells, and the `panel_rmse()` placebo runs are run in parallel
  when `workers > 1` (using `future.apply`/`future`, if installed). Estimates
  remain reproducible given `seed`.

```{r perf, eval = FALSE}
trop(df, "y", "w", "id", "t", se = "bootstrap",
     control = trop_control(svd = "truncated", workers = 4))
```

## Reference

Athey, S., Imbens, G. W., Qu, Z., & Viviano, D. (2025). *Triply Robust Panel
Estimators.* arXiv:2508.21536.
